---
title: "Sectoral Employment Dynamics in Australia"
author:
- familyname: Yang (ID:30306396)
  othernames: (Elvis) Zhixiang 
  address: Monash University
  email: zyan0056@student.monash.edu
  correspondingauthor: true
  qualifications: \textbf{Email :zyan0056@student.monash.edu}
- familyname: Vahid
  othernames: Farshid
  address: Monash University
  email: farshid.vahid@monash.edu
  correspondingauthor: true
  qualifications: \textbf{Email :farshid.vahid@monash.edu}
department: Department of\newline Econometrics &\newline Business Statistics
organization: \textbf{EBS Honours Research Plan}, supervised by \textbf{Farshid Vahid}
bibliography: references.bib
biblio-style: apa
linestretch: 1.5
output:
  monash::report:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
library(ggplot2)
library(tidyverse)
```

\graphicspath{ {/Users/elvisyang/Desktop/Sectoral Employment Forecsating /Research_Proporsal/Elvis_proporsal/figure} }

# Introduction

The COVID-19 pandemic has had a massive effect on economies around the world. Across different countries, millions of workers were furloughed or even lost their jobs as businesses struggled to survive [@ny2020]. Same situation happened in Australia, due to more restrictions, many businesses closed their doors, while employees were working with less hours or being dismissed by companies. As a result of the continuous "lockdown" periods in 2020, the estimates made by the Australian Bureau of Statistics [@ABS2021] concluded that 72% of business generated less revenue and the underemployment rate hit historically high with 13.8% by the end of April, 2020, only one month after the COVID-19 outbreak.

Our research is motivated by the lack of quantitative research on the employment of subsectors in Australia, as many studies have focused on the aggregated employment rate. A general problem of aggregated researches is the loss of hierarchical information, which may result in a biased conclusion or "a illusion of employment prosperity". Thus, a quantitative analysis of the sectoral employment will overcome this problem, giving us a better scope to evaluate the impacts of COVID-19 in Australia.


# Research aim and questions

This research will extend @anderson2020 by using data on 87 subsectors, instead of 19 that they used. I will develop a model for the subsectors to evaluate the long run effect and the COVID-19 post-impacts. I will also provide an contrafactual analysis based on the assumption "if there is no pandemic". Subsectoral data will provide us more information, which will assist in getting a better understanding of employment dynamics in Australia.

The overall research aim is to provide estimates of subsectoral employment based on historical data. Specifically, my goals are: 

1. To construct a time series model of employment in 87 subsectors of the Australian economy.

1. To use this model to conduct our contrafactual analysis. 

3. To use this model to determine which sectors have the highest impact (or positive spillover) on employment growth in the long run.


\newpage

# Review of literature

Our review of literature mainly focuses on two areas:

1.  The COVID-19 sectoral impacts and modelling to Economies.

2.  Modelling of large numbers of time series.

## Sectoral Impact for COVID-19 to Economies

Most existing studies have focused on the evaluation of impacts of COVID-19 on broad sectors of large economies such as the US and Europe. @ludvigson2020covid developed a disaster series to translate the macroeconomic impact of costly and deadly disasters in recent US history and modelled as sectoral shock to predict COVID-19, concluding that the shock would lead to a cumulative loss of 20% in industrial production, 39% in public services and also reduce the US GDP by 12.75 per cent at the end of 2020. @gregory2020pandemic conducted simulations under different scenarios via a search theoretic model using US data and found the recovery in US is L-shaped, with employment remaining lower than the pre-covid for a long period. They also extended their studies at disaggregated level of 20 sectors, showing that "arts and entertainment" and "accommodation and food services" sectors would have the biggest shock during the pandemic.

<!-- @berger2020nowcasting also mentioned that a fall of 10.7 percent of output gap for US in third quarter.  -->

In Australia, @anderson2020 developed a multivariate time series for 19 main sectors in Australia (as a small open economy) using a Bayesian VARX model. Their research concluded that "Manufacturing" and "Construction" have highest positive spillovers for the aggregate economy. Meanwhile, they also applied a "conditional forecasting" method proposed by @waggoner1999 to simulate different scenarios for pandemic in Australia. However, this research does not use the most disaggregted level in Australia (subsectors of main sectors), which can be extremely useful in macroeconomic analysis.

<!-- should I put the part 1 last part here ??? for @anderson2020 or keep the existing one -->

## Modelling

### Baysian VAR

Literature of multivariate, large-size data tend to rely on Bayesian Vector Autoregression model(BVAR) [e.g. @anderson2020; @litterman1986; @banbura2010large]. The BVAR model is attractive because it allows us to estimate a large number of parameters, when sample size is not large, in a statistically coherent way.[@litterman1986;@wozniak2016bayesian].

In order to utilize the Bayesian VAR estimators and decrease the weight of the lagged variable with the lag length, @litterman1979 proposed the Minnesota Prior. The prior mean on the first own lag is set to unity and the rest are set to zero so that (a) the most recent lag should provide more information than distant lags. (b) own lags should explain more than the lags of other variables.

### Improvement of BVAR

The literature suggests that a significant improvement can be made in the large BVAR dynamic model by imposing a stronger shrinkage parameter [@banbura2010large; @litterman1986]. @robertson1999vector and @kadiyala1997 proposed a normal inverted Wishart prior which retains the principal of Minnesota prior. Meanwhile, @banbura2010large suggested an easier way to apply the Minnesota prior via adding dummy observations in the BVAR system.

# Data collection and exploratory analysis

## Data Introduction

<!-- Find out some points about the data especially focusing on 19 to 87 comparison , identify the advantage of using 87 sectors rather than 19 sectors  -->

Our data source is the ABS Employment by industry subdivision of main job [@ABS2022], which records employment (measured in $\bf{'000}$) by the ANZSIC industry sub-division of their main jobs from $1984:Q4$ to $2021:Q4$. Our data structure is provided via Figure \ref{fig:anzsic}.

Although seasonally adjusted data is available in [@ABS2022], I will work with original data to capture any possible changes in seasonal patterns. An seasonally difference will be applied to the logarithm of the original series in order to make it stationary and eliminate seasonality, which will make it easier to conduct further steps.

## Priliminary Exploratory Data Analysis

\newpage

Figure \ref{fig:19} illustrates the changes of raw data for 19 sectors in a more disaggregated manner. Due to the closedown of businesses and travel bans on 2020:Q2, we can observe that the number dropped substantially (from around 13200 to 12200 on $2020:Q2$). Most industries behaved similarly with significant changes shown in Figure \ref{fig:19} . Comparing with the previous data of these industries," Accommodation & Food", "Media & telecom" and Administrative industries have experienced a severe loss of employment and have not fully recovered to the pre-covid level. However, some industries like "Financial" and "Electricity & Gas" show an continuously increase trend as before.

Meanwhile, at 87 subsector level (the most disaggregated) in Table \ref{tab:comp}, we can conclude that the Food and Beverage Service experienced a severe shock after the lockdown, following by Heritage Activities and Other Store-Based Retailing.Figure \ref{fig:87} demonstrates the performance of each industry in the most disaggregated level, we can see subsectors we mentioned before have shown huge decreases in employment. It further proves that the employment decreases hugely in those subsectors mentioned before.

Nevertheless, there is a draw back of using 19 sectoral employment, that is subsectoral dynamics of these sectors may not be homogeneous with their aggregated sectoral changes. For example, when observing the aggregated performance of "Manufacturing" and "Mining" sectors from the 19 sectoral level (see Figure \ref{fig:a19}), we may believe that their corresponding subsetors should illustrate the same pattern. However, the reality is that while there is a decreasing trend in "Manufacturing" sector or an increasing trend in "Mining" sector, some of their subsectors are performed differently (see Figure \ref{fig:a87}). This means that not all subsectors follow the same pattern with the aggregated sectoral level.


```{=tex}
\begin{table}
\begin{center}
\begin{tabular}{ccc}
\hline
Date     & Sector                                        & Percentage Changes \\
\hline
2020:Q2  & 27 Gas Supply                                 & 50.48\%                                \\
2020: Q2 & 57 Internet Publishing and Broadcasting       & 44.44\%                                \\
2020: Q2 & 56 Broadcasting                               & 41.43\%                                \\
2020: Q2 & 26 Electricity Supply                         & 36.34\%                                \\
2020: Q2 & 20 Non-Metallic Mineral Product Manufacturing & 32.37\% 
                 \\
                 \\
                 \hline
Date     &  Sectors                                      &Percentage Changes\\
                 \hline
2020: Q2 & 03 Forestry and Logging                   &-55.88\%  
                 \\
2020: Q2 & 96 Private Households Employing Staff     & -49.09\% 
                 \\
2020: Q2 & 60 Library and Other Information Services & -46.60\% 
                 \\
2020: Q2 & 02 Aquaculture                                & -44.58\%  
                 \\
2020: Q2 & 91 Sports and Recreation Activities           & -43.71\%
\end{tabular}
\end{center}
\caption{The highest and lowest five subsectors' employment percentage change for 2020:Q1 to 2020:Q2 }
\label{tab:comp}
\end{table}
```

# Methdology
## Proposed Model

I plan to use a Bayesian VARX model based on a method proposed by @anderson2020. In the model, each sector is affected by the lags of sectoral growth and a lag of the total employment growth. The lag of aggregate employment growth act as an economy-wide factor which will affect each sector.

Under the assumption that the structure of Australian economy will not change during the COVID-19, we suggest the BVAR model as

$$
\begin{aligned}
\textbf{y}_t=\textbf{c}+\textbf{A}_1 \textbf{y}_{t-1}+\textbf{A}_2\textbf{y}_{t-2}+\textbf{A}_3\textbf{y}_{t-3}+\textbf{A}_4\textbf{y}_{t-4}+\bf{\Gamma}\textbf{x}_{t-1}+\bf{u}_t
\end{aligned}
$$

where $\bf{y}_t$ is an $87\times1$ vector of subsectoral employment growth rate at time $t$  and $\bf{x}_{t-1}$ is the $4\times1$ vector of 4 lags of the growth rate of aggregate employment (this vector of variables are predetermined at time $t$), **c** is a vector of constants, $\bf{A}_{1,2,3,4}$ are $87\times87$ parameter matrices. $\bf{\Gamma}$ is a $87\times4$ matrix and $\bf{u}_t$ is a vector of reduced form errors with the mean equals to zero and independent variance $\bf{u}_t \sim (0,\bf{\Sigma})$. (see Appendix)

## Prior and shrinkage

I plan to estimate the VARX using Bayesian methods by specifying a Minnesota prior [e.g. @anderson2020; @litterman1986; @robertson1999vector]. In order to set up the Minnesota prior in our BVAR model, @banbura2010large suggestes a nature conjugate Normal-Wishart prior that applies shrinkage to the VAR slope coefficients using a Minnesota-type prior as follows:

$$
\begin{aligned}\label{eq:1}
&E[a_{j}^{jk}] = E[\gamma_{i}^j]=0\\
\\
&Var[a_j^{jk}]= 
\begin{cases}
\frac{\lambda^2}{i^2},&j=k\\
\frac{\lambda^2}{i^2}\frac{\sigma^2_{j}}{\sigma^2_k},& otherwise
\end{cases}\\
\\
&Var[\gamma_i^{j}]=\frac{\lambda^2}{i^2}\frac{\sigma^2_{j}}{\sigma^2_e}
\end{aligned}
$$
where the degree of shrinkage is governed by $\lambda$, $\frac{1}{i^2}$ to down-weight more distant lags and the $\frac{\sigma_j^2}{\sigma_k^2}$ adjust for different scales. 

@banbura2010large stated that the natural conjugate Normal-Inverse-Wishart posterior moments can be calculated either analytically or through adding the dummy observations. I will use dummy observations to estimate the BVAR e.g. [@banbura2010large; @anderson2020; @wozniak2016bayesian]. More details are provided in the Appendix.

## Sectoral Employment Multiplier Analysis 

I will also conduct an subsectoral employment multiplier analysis based on the estimated BVAR model and the 87 subsectoral data. 

The current economy multiplier is : 

$$
Total\ Employment=\sum_{j=1}^{87} w_j\times {GR}_j
$$

where $w_j$ is the share of employment of subsector $j$ in the total employment and $GR$ is the growth rate in employment of subsector $j$. 

Due to the interconnection of macroeconomic subsectors, when a subsector $j$ has an increase in employment, in the long run, it may have spillover effect onto other subsectors, especially those that have high connections with the subsector $j$ (@anderson2020).

If this long-run effect is larger than the immediate effect ($w_j\times GR_j$), then this subsector will have positive spillover effect on total employment, *vice versa*. I will use the interconnection of BVAR parameters to estimate subsectors which have strong positive spillovers for the total employment. Then, if the government makes policies to stimulate these subsectors, the total employment will recover more effectively.  


# Timeline and expected outcome

Table \ref{tab:timeline1} summarises the work done to date. Table \ref{tab:timeline2} maps out the plan for completing the thesis research.

```{r timeline, results='asis'}
timeline = data.frame(Timeline = c("Week 2","Week 3","", "Week 4","", "Week 5", "Week 6", "Week 7","Week 8", "week 9"),
                      Tasks = c(" Background reading about Multivariate VAR model",
                                "Collect subsectoral data from ABS website",
                                "Prepare materials reltaed to subsectoral employment evaluations",
                                "Prepare literature review related to BVAR model and setting priors",
                                "Analysis about the MATLAB code done by Andereson et al. (2020)",
                                "Read articles about Conditional Forecasting",
                                "Prepare the conditional forecasting coding and hierarchical forecasting",
                                "Exploratory data analysis on sectoral and subsectoral data",
                                "Select suitable plots and test softwares to prepare the proposal ",
                                "Write research proposal and prepare the first presentation"))


knitr::kable(timeline, "latex",
             caption = "Completed work",
             label = "timeline1",
             booktabs = TRUE,
             escape = FALSE)
  
```

```{r}
timeline = data.frame(Timeline = c("June - July", "August", "September", "October"),
                      Tasks = c("Modelling of subsectoral employment", "Train the model using pre-covid data with BVAR model and compare the accuracy with other models","apply forecasts using BVAR for quarters between 2020 to 2022, conduct both conterfactual analysis and sectoral multiplier analysis", "Write my thesis and prepare for the second presentation")) 

knitr::kable(timeline, "latex",
             caption = "Research plan",
             label = "timeline2",
             booktabs = TRUE,
             escape = FALSE) 
```

\newpage

# Acknowledgement

I would like to gratitude my supervisor Professor Farshid Vahid and my coordinator Professor Heather Anderson very much for their selfless support and devoted care along the way.

# Appendix

The VARX model is: 

```{=tex}
\begin{align}
\textbf{y}_t&=\textbf{c}+\textbf{A}_1 \textbf{y}_{t-1}+\textbf{A}_2+\cdots+\textbf{A}_4\textbf{y}_{t-4}+\bf{\Gamma}_1\textbf{x}_{t-1}+\bf{u}_t\\
&=
\begin{bmatrix}
c_1\\
\vdots\\
c_2
\end{bmatrix}
+
\begin{bmatrix}
a_1^{11}&\cdots&a_1^{1n}&\cdots&a_4^{11}&\cdots&a_4^{1n}&\gamma_1^{1}&\cdots&\gamma_4^{1}\\
\vdots&\ddots&\vdots&\ddots&\vdots&\ddots&\vdots&\vdots&\ddots&\vdots\\
a_1^{n1}&\cdots&a_1^{nn}&\cdots&a_4^{n1}&\cdots&a_4^{nn}&\gamma_1^n&\cdots&\gamma_4^n\\
\end{bmatrix}
\begin{bmatrix}
\bf{y}_{t-1}\\
\vdots\\
\bf{y}_{t-4}\\
x_{t-1}\\
\vdots\\
x_{t-4}
\end{bmatrix}\\
&+
\begin{bmatrix}
u_{1,t}\\
\vdots\\
u_{n,t}
\end{bmatrix}\\
\end{align}
```
where $E(\bf{u}_t\bf{u}'_t)=\bf{\Sigma}$ and $E(\bf{u}_t\bf{u'}_{t-1})=0$. Here the $n$ represent the number of sectors (or in our case will be 87) and $c$ represents the vector of constants. There are 4 lags included for the total employment $(x_{t-1},\cdots,x_{t-4})$ as predetermined variable at time t.

Here is an Minnesota prior example done by @anderson2020:

The Normal-Wishart prior distribution take the form of

$$
\begin{aligned}
&vec(\mathbf{\beta}|\Sigma)\sim N(vec(\bf{\beta_0}),\Sigma\otimes\Omega_0)\ and\\
&\Sigma\sim\mathbf{IW}(S_0,a_0)
\end{aligned}
$$ 

where we set the prior parameters $\bf{\beta_0,\Omega_0,S_0}$ and $a_0$ based on the prior setting in \ref{eq:1}. The expectation of $\Sigma$ being $diag(\sigma^2_1,\cdots,\sigma^2_n)$. 


Then we implement our VAR by defining dummy observations. 

$$
\begin{aligned}
Y_d&=
\begin{pmatrix}
\bf0_{np+p,n}\\
\bf diag(\bf{\sigma_1,\cdots,\sigma_n})\\
\bf0_{1\times n}
\end{pmatrix}\\
X_d&=
\begin{pmatrix}
\bf{J_p}\otimes diag(\frac{\sigma_1}{\lambda}\cdots\frac{\sigma_n}{\lambda},\frac{\sigma_e}{\lambda})&\bf0_{(np+p)\times1}\\
\bf 0_{n,np+p}&\bf 0_{n\times1}
\\
\bf 0_{1,np+p}&\bf \epsilon
\end{pmatrix}
\end{aligned}
$$ where

$$
\begin{aligned}
\bf{J_p}=diag(1,\cdots,p)\\
\bf{S_0}=(Y_d-X_d\times B_0)'(Y_d-X_dB_0)\\
\bf{B_0}=(X_d'X_d)^{-1}X_dY_d,\ \Omega_0=(X_d'X_d)^{-1}\  and\\
a_0=T_d-np-p-1\\
\end{aligned}
$$ 

where $T_d$ is the number of rows for both $\bf{Y}_d$ and $\bf{X}_d$.

We can get 
\begin{align}
\bf{Y^*}=\bf{X^*}\bf{\beta}+\bf{\mu}^*\ \ \  where: \\
\bf{Y^*}=[\bf{Y'},\bf{Y'_d}]';\ \bf{X^*}=[\bf{X'},\bf{X'_d}]';\ \bf{\mu^*}=[\bf{\mu'},\bf{\mu'_d}]'
\end{align}


Then we can estimating the BVAR by conducting an least square regression of $Y^*$ on$X^*$. The posterior distribution then have the form of

\begin{align}
&vec(\mathbf{\beta})|\Sigma,Y\sim N(vec(\bf{\tilde\beta}),\Sigma\otimes(\bf{X^*}'\bf{X^*})^{-1})\ and\\
&\Sigma|Y\sim\mathbf{IW}(\tilde\Sigma,T_d+T-np+2)
\end{align}


where $\bf{\tilde\beta} =(X*'X*)^{-1}X^{*'}Y^*$ and $\tilde\Sigma=(\bf{Y^*}-\bf{X^*}\bf{\tilde\beta})'(\bf{Y^*}-\bf{X^*}\bf{\tilde\beta})$




\newpage

\begin{figure}[t]
\includegraphics[scale=0.6]{free_19}
\centering
\caption{Employment('000) of 19 sectors in Australia from 2019:Q1 to 2021:Q4}
\label{fig:19}
\end{figure}


\begin{figure}[t]
\includegraphics[scale=0.5]{free_87}
\centering
\caption{Employment('000) of 87 subsectors in Australia from 2019:Q1 to 2021:Q4}
\label{fig:87}
\end{figure}


\begin{figure}[t]
\includegraphics[scale=0.5]{agg_19}
\centering
\caption{Aggregated Employment(in thousands) for Manufacturing and Mining sector in Australia from 1984:Q4 to 2021:Q4}
\label{fig:a19}
\end{figure}

\begin{figure}[t]
\includegraphics[scale=0.5]{agg_87}
\centering
\caption{Disaaggregate Employment(in thousands) of 87 subsectors in Manufacturing and Mining sector from 1984:Q4 to 2021:Q4}
\label{fig:a87}
\end{figure}

\begin{figure}[t]
\includegraphics[scale=0.5]{ANZSIC}
\centering
\caption{Australian Industry Pamamid plot by (ANZSIC)}
\label{fig:anzsic}
\end{figure}

\clearpage
